\documentclass[11pt]{fsuthesis}
%\usepackage[overload]{textcase}
%\usepackage[dvips]{graphics}
\usepackage{hyperref}
%\hypersetup{breaklinks=true}
%\usepackage{natbib}
\usepackage{setspace}

\usepackage[numbers,sort&compress]{natbib}

\usepackage{verbatim}
\usepackage{amsmath, amssymb, graphics}
\usepackage[pdftex]{graphicx}
\usepackage{subfig}
\usepackage[chapter]{algorithm}
\usepackage{algorithmic}


\input{macros.tex}
\input{misc_mac.tex}

\title{FLOCKING IMPLEMENTATION FOR THE BLENDER GAME ENGINE}
\author{Myrna I. Merced Serrano}
\college{College of Arts and Sciences}
\department{Department of Scientific Computing}
\manuscripttype{Thesis}
\degree{Master of Science}
\semester{Summer}
\degreeyear{2011}
\defensedate{June 24, 2011}

\committeesig{Gordon Erlebacher}{Professor Directing Thesis}
%\committeesig{XXX}{University Representative}
\committeesig{Ming Ye}{Committee Member}
\committeesig{Xiaoqiang Wang}{Committee Member}

\collegesig{Max Gunzburger}{Chair}{Department of Scientific Computing}
\collegesig{Joseph Travis}{Dean}{College of Arts and Sciences}

\begin{document}

\frontmatter
\maketitle
\makesignaturepage

\begin{dedication}
The author of this thesis wants to dedicate it to her family for their support.
\end{dedication}

\begin{acknowledgments}
The author of this thesis would like to acknowledge the unconditional help of her advisor Dr. Gordon Erlebacher, his support and advice helped her to improve her knowledge and motivated her to work harder everyday. Also, she would like to thanks her co-worker Ian Johnson for his availability to help her and for developing most of the code used in this thesis. Last but not least, she would like to thanks her DSC graduate fellows, faculty and staff for their support.
\end{acknowledgments}

\tableofcontents
\listoftables
\listoffigures
%\listofmusex

%\begin{listofsymbols}
%\end{listofsymbols}

%\begin{listofabbrevs}

%\begin{acronym}{GPU}
%\acro{GPU}{Graphics Processor Units}
%\end{acronym}


%\end{listofabbrevs}

% spacing 1.5
\onehalfspacing

\begin{abstract}
%Blender is a free modeling/simulation software available since 1993. 
%It has recently been extended to include modeling, texturing, 
%animation, particle simulation, rendering, game creation, etc. 
%The Blender Game Engine is a very powerful tool, allowing simple 
%games to be created without the need for explicit programming. 
%Although Blender has extensive particle-based tools, including hair 
%styling, these are absent from the game module. A submodule of the 
%particle system is a rather sophisticated Boid System. 
In this thesis, we discuss the development of a new Boid 
system that simulates flocking behavior, inside the Blender 
Game Engine, and within the framework of 
the Real-Time Particles System (RTPS) library developed by Ian Johnson. 
%the capabilities of Blender. 
The collective behavior of Boids is characterized as an emergent behavior 
caused by following three steering behaviors: separation, alignment, 
and cohesion. 
%Boids are commonly used in games as non-player characters 
%since they can behave like real entities by themselves without the need 
%for explicit control. 
%We developed a new modifier for blender
%This modifier is called RTPS because it 
%depends on the library Real-Time Particles System (RTPS) developed by 
%Ian Johnson as part of his work on Smooth Particle Hydrodynamics (SPH). 
The implementation leverages the lastest OpenCL to maintain the portability 
of the Blender across different operating systems. 
%RTPS is a library that currently defines three particle systems: SPH, 
%FLOCK, and Simple. RTPS also incorporates CPU and GPU implementations 
%of both systems. 
Benchmarks of the RTPS-FLOCK system show that 
our implementation speeds up Blender's original Boid implementation (which 
cannot run in the game engine) by more than an order of magnitude. 
%as better performance than the Blender Boids system 
%(available outside the game engine). 
We illustrate our boid system with three demonstrations. 
The first demo shows the symmetry that each steering behavior 
maintains. The second demonstration illustrates the behavior of 
a  swarm of bees approaching their hive, and the last demo 
illustrates a simulation of a crowd crowd constrained to 
a two-dimensional plane. 
\end{abstract}

\mainmatter

\input chapter1 		% Introduction
\input chapter2		% Related Work
\input chapter3		% Flocking Algorithm
\input chapter4		% RTPS Library
\input chapter5		% Interface Developed for the Blender Game Engine
\input chapter6		% Interaction between Boids and SPH Particles
\input chapter7		% Conclusions
\input chapter8		% Future work

\appendix
\input appendix1

% You have your choice of bibliography sections, either
% hand-crafted or BibTeX.

% This is the "hand-crafted" bibliography/references section:
%\begin{references}
%Mybib, Sample.  \textit{An Example of a Bibligraphic Entry
% Created Manually}.  Tallahassee, Florida: Fornish and Frak, 2010.
%
%Smith, Marigold.  \textit{Lots and Lots of Bibliographic Entries
% and How to Display Them}.  Tallahassee, Florida: Gibson and Goulash, 2010.
%\end{references}

% The BibTeX bibliography/references section.  View the file
% 'myrefs.bib' to get a feel for what these entries may look like.

%\bibliographystyle{ieeetr} %plainnat
\bibliographystyle{unsrtnat}
\bibliography{myrefs}

\begin{biosketch}
Myrna Ivelisse Merced Serrano was born and raise at the southeast of the beautiful island of Puerto Rico. Myrna is an athletic girl, her favorite sport is basketball and at her free time she like to go out with her friends, cook, and watch television. 

She did her Bachelors degree in Computational Mathematics at the University of Puerto Rico at Humacao. While during her bachelors degree she became interesting on computers and programming. She went beyond most of the students of her class by gaining interest in research. She wanted to mix nanotechnology with biology, somehow. Therefore, she did studies on molecular dynamic simulations between carbon nanotubes and single-stranded DNA. Doing research motivates her to go farther from her current goal, that is why she decided to do graduate studies in a college in the United States. Although, she had to move far from her family to achieve her goals.

In August 2009 she joined the Department of Scientific Computing (DSC) at Florida State University. While she was there she worked as a teaching assistant, and she says that she enjoys the satisfaction of helping other students to enhance their knowledge. Also, the research presented in this thesis was done while she was a graduate student at DSC.
\end{biosketch}

\end{document}
